Prediction of 90-day mortality after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden, England and Wales

### 2020-06-11 reports/article.Rmd
revert to normal decimal points. Fix #.

```{r setup, include = FALSE}
options(
  digits            = 2,
  OutDec            = "\xB7",
  knitr.kable.NA    = '',
  repos = list(CRAN = "https://cran.rstudio.com/")
)
knitr::opts_chunk$set(
  echo = FALSE,
  out.width  = "100%", 
  out.height = "100%"
)
library(tidyverse)

@@

```{r}
# Change to ugly decimal point as required by LR
ud <- function(x) gsub("\\.", "\xB7", x)

shar_tot <- nrow(df)
shar_female <- df$P_Sex == "Female"
age_shar <- sprintf("%.0f [%.0f - %.0f]", mean(df$P_Age), min(df$P_Age), max(df$P_Age)) %>% ud()
sex_shar <- 
  sprintf(
    "%.0f %% [%s of %s]", 
    mean(shar_female) * 100, 
    format(sum(shar_female), big.mark = ",", decimal.mark = "\xB7"), 
    format(shar_tot, big.mark = ",", decimal.mark = "\xB7")
) %>% ud()
deaths_shar <- 
  sprintf(
    "%.02f %% [%.0f of %s]", 
    mean(df$death90) * 100, 
    sum(df$death90), 
    format(nrow(df), big.mark = ",")
  ) %>% ud()

shar_summary <- sprintf(
  "(mean [range] age %s; %s females; %s deaths within 90 days)",
  age_shar, sex_shar, deaths_shar
) %>% ud()

extr <- function(col, reg) 
  subset(table1_combined_tot, startsWith(what, col), reg, drop = TRUE)
extr_perc <- function(...) 
  gsub("([\\d,]*)( )(\\()(\\d*)(\\.\\d*)(\\))", "\\4 \\% \\[\\1 of", extr(...), perl = TRUE) %>% ud()

njr_age_mean <- as.numeric(gsub("(.*) (.*)", "\\1", extr("Age", "NJR"), perl = TRUE))
# Same range by design (to match Swedish cohort)
njr_age <- sprintf("%.0f [%.0f - %.0f]", njr_age_mean, min(df$P_Age), max(df$P_Age)) %>% ud()
njr_sex <- sprintf("%s %s]", extr_perc("Sex", "NJR"), format(N_njr$tot, big.mark = ",")) %>% ud()
njr_deaths <- 
  sprintf(
    "%.02f %% [%.0f of %s]", 
    (N_njr$death / N_njr$tot) * 100, 
    N_njr$death, 
    format(N_njr$tot, big.mark = ",")
  ) %>% ud()
njr_summary <- sprintf(
  "(mean [range] age %s; %s females; %s deaths within 90 days)",
  njr_age, njr_sex, njr_deaths
) %>% ud()

get_auc <- function(mod) {
  subset(
    brlasso_tbl_auc, 
    Model == mod & 
      `Age as` %in% c(NA, "main effect"), 
    "AUC", 
    drop = TRUE
  ) %>% 
  {sprintf("(AUC = %s", gsub(" (", ", 95 % CI ",., fixed = TRUE))} %>% 
  ud()
}

@@

`r format(nrow(df), big.mark = ",")` patients (mean age `r mean(df$P_Age)` years, range `r paste(range(df$P_Age), collapse = " - ")` and `r round(mean(df$P_Sex == "Female") * 100)` % females) were included in the derivation cohort from SHAR (Figure \@ref(fig:flowchart) left panel). `r with(df, sprintf("%d (%.2f %%)", sum(death90), mean(death90) * 100)) %>% ud()` of those patients died within 90 days and no one was censored before that. The unadjusted risk of 90-day mortality was therefore `r ud(surv90d)`. Further characteristics of the study population are presented in Table \@ref(tab:tab1). `r round(mean(df$CCI_index_quan_original != 0) * 100)` % and `r round(mean(df$ECI_index_sum_all  != 0) * 100)` % of the Swedish patients had at least one pre-surgery co-morbidity according to Charlson and Elixhauser. The proportion of patients with ASA class III was `r round(mean(df$P_ASA == 3) * 100)` %.

In addition, `r format(N_njr$tot, big.mark = ",")` patients were included for the external validation cohort from England and Wales (Figure \@ref(fig:flowchart) right panel). Their unadjusted risk of 90 day mortality was `r ud(njr_surv)`, which was higher compared to the derivation cohort from Sweden.

@@

Obesity was not statistically significant by itself ($p =$ `r subset(brlasso_tbl_coefs, term == "Obesity", "p", drop = TRUE) %>% ud()`) but was still relevant as a predictor due to unobserved heterogeneity. We noted that the proportion of patients registered in the SHAR with BMI above 30, the WHO definition of obesity, was much higher (`r round(mean(df$P_BMI > 30, na.rm = TRUE) * 100)` %, n = `r format(sum(df$P_BMI > 30 & !is.na(df$P_BMI)), big.mark = ",")`) than the proportion of patients with a diagnosis code for obesity (ICD-10 = E66) recorded in the Swedish national patient register (`r round(mean(df$c_obesity) * 100, 1)` %, , n = `r format(sum(df$c_obesity & !is.na(df$c_obesity)), big.mark = ",")`). We believe this might be due to the fact that BMI is systematically recorded in the SHAR, but the diagnosis of obesity might not have been made during previous contacts with the healthcare system.

>>>>

```{r setup, include = FALSE}
options(
  digits            = 2,
  knitr.kable.NA    = '',
  repos = list(CRAN = "https://cran.rstudio.com/")
)
knitr::opts_chunk$set(
  echo = FALSE,
  out.width  = "100%", 
  out.height = "100%"
)
library(tidyverse)

@@

```{r}
shar_tot <- nrow(df)
shar_female <- df$P_Sex == "Female"
age_shar <- sprintf("%.0f [%.0f - %.0f]", mean(df$P_Age), min(df$P_Age), max(df$P_Age))
sex_shar <- 
  sprintf(
    "%.0f %% [%s of %s]", 
    mean(shar_female) * 100, 
    format(sum(shar_female), big.mark = ","), 
    format(shar_tot, big.mark = ",")
)
deaths_shar <- 
  sprintf(
    "%.02f %% [%.0f of %s]", 
    mean(df$death90) * 100, 
    sum(df$death90), 
    format(nrow(df), big.mark = ",")
  )

shar_summary <- sprintf(
  "(mean [range] age %s; %s females; %s deaths within 90 days)",
  age_shar, sex_shar, deaths_shar
)

extr <- function(col, reg) 
  subset(table1_combined_tot, startsWith(what, col), reg, drop = TRUE)
extr_perc <- function(...) 
  gsub("([\\d,]*)( )(\\()(\\d*)(\\.\\d*)(\\))", "\\4 \\% \\[\\1 of", extr(...), perl = TRUE)

njr_age_mean <- as.numeric(gsub("(.*) (.*)", "\\1", extr("Age", "NJR"), perl = TRUE))
# Same range by design (to match Swedish cohort)
njr_age <- sprintf("%.0f [%.0f - %.0f]", njr_age_mean, min(df$P_Age), max(df$P_Age))
njr_sex <- sprintf("%s %s]", extr_perc("Sex", "NJR"), format(N_njr$tot, big.mark = ","))
njr_deaths <- 
  sprintf(
    "%.02f %% [%.0f of %s]", 
    (N_njr$death / N_njr$tot) * 100, 
    N_njr$death, 
    format(N_njr$tot, big.mark = ",")
  )
njr_summary <- sprintf(
  "(mean [range] age %s; %s females; %s deaths within 90 days)",
  njr_age, njr_sex, njr_deaths
)

get_auc <- function(mod) {
  subset(
    brlasso_tbl_auc, 
    Model == mod & 
      `Age as` %in% c(NA, "main effect"), 
    "AUC", 
    drop = TRUE
  ) %>% 
  {sprintf("(AUC = %s", gsub(" (", ", 95 % CI ",., fixed = TRUE))}
}

@@

`r format(nrow(df), big.mark = ",")` patients (mean age `r mean(df$P_Age)` years, range `r paste(range(df$P_Age), collapse = " - ")` and `r round(mean(df$P_Sex == "Female") * 100)` % females) were included in the derivation cohort from SHAR (Figure \@ref(fig:flowchart) left panel). `r with(df, sprintf("%d (%.2f %%)", sum(death90), mean(death90) * 100))` of those patients died within 90 days and no one was censored before that. The unadjusted risk of 90-day mortality was therefore `r surv90d`. Further characteristics of the study population are presented in Table \@ref(tab:tab1). `r round(mean(df$CCI_index_quan_original != 0) * 100)` % and `r round(mean(df$ECI_index_sum_all  != 0) * 100)` % of the Swedish patients had at least one pre-surgery co-morbidity according to Charlson and Elixhauser. The proportion of patients with ASA class III was `r round(mean(df$P_ASA == 3) * 100)` %.

In addition, `r format(N_njr$tot, big.mark = ",")` patients were included for the external validation cohort from England and Wales (Figure \@ref(fig:flowchart) right panel). Their unadjusted risk of 90 day mortality was `r njr_surv`, which was higher compared to the derivation cohort from Sweden.

@@

Obesity was not statistically significant by itself ($p =$ `r subset(brlasso_tbl_coefs, term == "Obesity", "p", drop = TRUE)`) but was still relevant as a predictor due to unobserved heterogeneity. We noted that the proportion of patients registered in the SHAR with BMI above 30, the WHO definition of obesity, was much higher (`r round(mean(df$P_BMI > 30, na.rm = TRUE) * 100)` %, n = `r format(sum(df$P_BMI > 30 & !is.na(df$P_BMI)), big.mark = ",")`) than the proportion of patients with a diagnosis code for obesity (ICD-10 = E66) recorded in the Swedish national patient register (`r round(mean(df$c_obesity) * 100, 1)` %, , n = `r format(sum(df$c_obesity & !is.na(df$c_obesity)), big.mark = ",")`). We believe this might be due to the fact that BMI is systematically recorded in the SHAR, but the diagnosis of obesity might not have been made during previous contacts with the healthcare system.

### 2020-06-11 reports/tabs.Rmd
revert to normal decimal points. Fix #.

```{r tab1}
knitr::kable(
  mutate(table1_combined_tot, what = gsub("Heart infarct", "Myocardial infarction", what)), 
  caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anesthesiologists classification. CNS = central nervous system.")
```

>>>>

```{r tab1}
table1_combined_tot %>% 
  mutate(what = gsub("Heart infarct", "Myocardial infarction", what)) %>% 
knitr::kable(
    caption = "Characteristics of the study population in the model derivation cohort (SHAR) and external validation cohort (NJR). Educational levels were classified as low (up to 9 years), middle (10-12 years) and high (at least 12 years).  BMI = Body mass index. ASA = American Society for Anesthesiologists classification. CNS = central nervous system.")
```
